---
title: "Metabolizer"
output: html_notebook
---
# 1st step

```{r}

setwd("G:/My Drive/data/CCLE and BRCA project from 2013/paper all figures March 2016/Cell Resource Paper 2020/Fig1_gene_expression/METABRIC/Data compilation/")


library(dplyr)
library(gplots)
library(tidyr)
library(tibble)
library(stringr)
library(MCbiclust)
library(gdata)
library(WriteXLS)
library(biomaRt)
library(pheatmap)
library(propagate)
library(GSVA)
library(openxlsx)

setwd("G:/My Drive/data/CCLE and BRCA project from 2013/paper all figures March 2016/Cell Resource Paper 2020/Fig1_gene_expression/METABRIC/Data compilation/")

load('METABRIC_DATA.RData')
load('METABRIC_gene_lists.RData')
load('METABRIC_sort_data.RData')
load('METABRIC_PC1_GSEA.RData')
```

# Step 2: get the correlation of each sample to the forks

```{r}
sort.df <- data.frame(sample = Complete_METABRIC_Clinical_Features_Data[,1],
                      MB3.index = order(mito.sort[[1]]),
                      MB3.pc1  = mito.pc1[[1]][order(mito.sort[[1]])],
                      MB2.index = order(mito.sort[[2]]),
                      MB2.pc1  = mito.pc1[[2]][order(mito.sort[[2]])],
                      Mito3.index = order(mito.sort[[3]]),
                      Mito3.pc1  = mito.pc1[[3]][order(mito.sort[[3]])],
                      MB1.index = order(ICT1.sort[[1]]),
                      MB1.pc1  = ICT1.pc1[[1]][order(ICT1.sort[[1]])],
                      Random1.index = order(random.sort[[1]]),
                      Random1.pc1  = random.pc1[[1]][order(random.sort[[1]])],
                      Random2.index = order(random.sort[[2]]),
                      Random2.pc1  = random.pc1[[2]][order(random.sort[[2]])])


all.clinical.df <- left_join(sort.df,
                             Complete_METABRIC_Clinical_Features_Data,
                             by = 'sample')
```

# top MB1 and MB2

```{r}
MB1.UF <- all.clinical.df %>%
  filter(MB1.pc1 > 0) %>%
  slice_min(MB1.index, n = 10)
MB1.UF

MB2.UF <- all.clinical.df %>%
  filter(MB1.pc1 > 0) %>%
  slice_min(MB2.index, n = 10)
MB2.UF

MB1.LF <- all.clinical.df %>%
  filter(MB1.pc1 < 0) %>%
  slice_min(MB1.index, n = 10)
MB1.LF

MB2.LF <- all.clinical.df %>%
  filter(MB1.pc1 < 0) %>%
  slice_min(MB2.index, n = 10)
MB2.LF
```

#Change .dots by -dashes 

```{r}
names(METABRIC.data) <- gsub('\\.','-',names(METABRIC.data))  
```

#Extract metabolic genes from METABRIC.data

```{r}
metab.genes.xls <- read.xlsx('GS metabolic genes list.xlsx', sheet=1)
metab.METABRIC <- filter(METABRIC.data, METABRIC.data[,1] %in% metab.genes.xls[,2])

```


# Tranform sample row into a list of samples, and then select those samples from METABRIC file 

```{r}
columns1L <- MB1.LF[,1]
columns1L <- append(columns1L, "external_gene_name", after = 0)
MB1.LF.samples <- metab.METABRIC[,columns1L]

columns1U <- MB1.UF[,1]
columns1U <- append(columns1U, "external_gene_name", after = 0)
MB1.UF.samples <- metab.METABRIC[,columns1U]

columns2L <- MB2.LF[,1]
columns2L <- append(columns2L, "external_gene_name", after = 0)
MB2.LF.samples <- metab.METABRIC[,columns2L]

columns2U <- MB2.UF[,1]
columns2U <- append(columns2U, "external_gene_name", after = 0)
MB2.UF.samples <- metab.METABRIC[,columns2U]
```

# Create Excel files from the samples selected previously 

```{r}
write.xlsx(MB1.UF.samples,
           file = 'MB1.UF.samples.xlsx', overwrite = TRUE)

write.xlsx(MB2.UF.samples,
           file = 'MB2.UF.samples.xlsx', overwrite = TRUE)

write.xlsx(MB2.LF.samples,
           file = 'MB2.LF.samples.xlsx', overwrite = TRUE)

write.xlsx(MB1.LF.samples,
           file = 'MB1.LF.samples.xlsx', overwrite = TRUE)

```

#After this script starts the annoying part, the one I could not optimize:
#In order to make the data run on Metabolizer: 

#->the external_gene_names need to converted to ENTREZ numeric codes using gprofiler

#->the data needs to be checked for repetitions because gprofiler creates them

#->at least 2 conditions (MB1.LF and MB2.LF, for example) need to be included in the same TSV file, after a numeric entrez id in the 1st column

#->There needs to be a seperate TSV file describing which sample belongs to which group

#->the process is dificult to explain, the best thing is to download the github file of metabolizer, which includes dummy data in the format required.