---
title: "Metabolizer"
output: html_notebook
---
# This is a script I did to be able to run our dataset in the Metabolizer script(metabolizer.babelomics.org/)

```{r}


setwd("C:/Users/Utente/Documents/Portfolio/R/")


library(dplyr)
library(gplots)
library(tidyr)
library(tibble)
library(stringr)
library(MCbiclust)
library(gdata)
library(WriteXLS)
library(biomaRt)
library(pheatmap)
library(propagate)
library(GSVA)
library(openxlsx)


load('METABRIC_DATA.RData')
load('METABRIC_gene_lists.RData')
load('METABRIC_sort_data.RData')
load('METABRIC_PC1_GSEA.RData')
```

# Step 2: get the correlation of each sample to the forks

```{r}
sort.df <- data.frame(sample = Complete_METABRIC_Clinical_Features_Data[,1],
                      MB3.index = order(mito.sort[[1]]),
                      MB3.pc1  = mito.pc1[[1]][order(mito.sort[[1]])],
                      MB2.index = order(mito.sort[[2]]),
                      MB2.pc1  = mito.pc1[[2]][order(mito.sort[[2]])],
                      Mito3.index = order(mito.sort[[3]]),
                      Mito3.pc1  = mito.pc1[[3]][order(mito.sort[[3]])],
                      MB1.index = order(ICT1.sort[[1]]),
                      MB1.pc1  = ICT1.pc1[[1]][order(ICT1.sort[[1]])],
                      Random1.index = order(random.sort[[1]]),
                      Random1.pc1  = random.pc1[[1]][order(random.sort[[1]])],
                      Random2.index = order(random.sort[[2]]),
                      Random2.pc1  = random.pc1[[2]][order(random.sort[[2]])])


all.clinical.df <- left_join(sort.df,
                             Complete_METABRIC_Clinical_Features_Data,
                             by = 'sample')
```

# top MB1 and MB2

```{r}
MB1.UF <- all.clinical.df %>%
  filter(MB1.pc1 > 0) %>%
  slice_min(MB1.index, n = 10)
MB1.UF

MB2.UF <- all.clinical.df %>%
  filter(MB1.pc1 > 0) %>%
  slice_min(MB2.index, n = 10)
MB2.UF

MB1.LF <- all.clinical.df %>%
  filter(MB1.pc1 < 0) %>%
  slice_min(MB1.index, n = 10)
MB1.LF

MB2.LF <- all.clinical.df %>%
  filter(MB1.pc1 < 0) %>%
  slice_min(MB2.index, n = 10)
MB2.LF
```

#Exchange .dots by -dashes 

```{r}
names(METABRIC.data) <- gsub('\\.','-',names(METABRIC.data))  
```

#Extract metabolic genes from METABRIC.data

```{r}
metab.genes.xls <- read.xlsx('metabolic genes list.xlsx', sheet=1)
metab.METABRIC <- filter(METABRIC.data, METABRIC.data[,1] %in% metab.genes.xls[,2])

```


# Tranform sample row into a list of samples, and then select those samples from METABRIC file 

```{r}
columns1L <- MB1.LF[,1]
columns1L <- append(columns1L, "external_gene_name", after = 0)
MB1.LF.samples <- metab.METABRIC[,columns1L]

columns1U <- MB1.UF[,1]
columns1U <- append(columns1U, "external_gene_name", after = 0)
MB1.UF.samples <- metab.METABRIC[,columns1U]

columns2L <- MB2.LF[,1]
columns2L <- append(columns2L, "external_gene_name", after = 0)
MB2.LF.samples <- metab.METABRIC[,columns2L]

columns2U <- MB2.UF[,1]
columns2U <- append(columns2U, "external_gene_name", after = 0)
MB2.UF.samples <- metab.METABRIC[,columns2U]
```

# Create Excel files from the samples selected previously 

```{r}
write.xlsx(MB1.UF.samples,
           file = 'MB1.UF.samples.xlsx', overwrite = TRUE)

write.xlsx(MB2.UF.samples,
           file = 'MB2.UF.samples.xlsx', overwrite = TRUE)

write.xlsx(MB2.LF.samples,
           file = 'MB2.LF.samples.xlsx', overwrite = TRUE)

write.xlsx(MB1.LF.samples,
           file = 'MB1.LF.samples.xlsx', overwrite = TRUE)

```

#the data obtained in this script then needs to go trought a manual step using gprofiler (biit.cs.ut.ee/gprofiler/gost), which converts the external_gene_names to ENTREZ numeric codes and controlled for repetitions.


#->There needs to be a seperate TSV file describing the experimental setup of each group

#->And then the data is ready to be run on Metabolizer (http://metabolizer.babelomics.org/)